### 2022-04-29 EDA_Chang_Noordermeer_2021.Rmd
first commit

---
title: "Analysis"
author: "Mikhail Dozmorov"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
```

# Libraries

```{r libraries}
library(tidyverse)
library(readxl)
library(writexl)
library(GenomicRanges)
```

# Settings

```{r settings}
# Project folder path
dir_data <- "/Users/mdozmorov/Documents/Work/GitHub/CTCF.dev"

# Input files
fileNameIn1 <- file.path(dir_data, "data/Chang_Noordermeer_2021.xlsx")
# Output files
fileNameOut1 <- file.path(dir_data, "")
```

# Load data

```{r data}
# Read in data
mtx <- read_xlsx(fileNameIn1, skip = 1)
```

## EDA

```{r}
summary(mtx$`-log10  p-value...7`)
summary(mtx$`-log10  q-value...8`)
summary(mtx$`-log10  p-value...13`)
summary(mtx$`-log10  q-value...14`)
table(mtx$`overlap with repeat(s)`)
sum
```

# GRanges

```{r}
gr <- GRanges(seqnames = mtx$coordinate, IRanges(start = mtx$...3, end = mtx$...4), strand = mtx$strand)
gr$ID <- mtx$number
gr$length <- mtx$length
gr$value <- mtx$value
gr$peak.log10pval <- mtx$`-log10  p-value...7`
gr$peak.log10qval <- mtx$`-log10  q-value...8`
gr$sequence <-mtx$sequence
gr$CTCF.log10pval <- mtx$`-log10  p-value...13`
gr$CTCF.log10qval <- mtx$`-log10  q-value...14`
gr$repeats <- mtx$`overlap with repeat(s)`

genome_id <- "mm10"
# Get chromosome info and match it to the chromosome order in ctcfBED
chrom_data <- GenomeInfoDb::getChromInfoFromUCSC(genome = genome_id)
chrom_data <- chrom_data[chrom_data$chrom %in% seqlevels(gr), ]
chrom_data <- chrom_data[match(seqlevels(gr), chrom_data$chrom), ]
# Check if chromosome order is the same
if (!all.equal(seqlevels(gr), chrom_data$chrom)) {
  print(paste("Chromosome order does not match for", genome_id, "genome."))
  break
}
# Assign seqinfo data
seqlengths(gr) <- chrom_data$size
isCircular(gr) <- chrom_data$circular
genome(gr)     <- genome_id

gr

# Save as RData object. subdir is the character name of the CTCF GRanges variable
fileNameOut <- "CTCF_mm10_Chang_Noordermeer_2021.RData"
save(list = gr, file = fileNameOut)
# load(file = paste0(subdir, ".RData"))

```

"ID" - numerical ID of a region with multiple (overlapping) CTCF ChIP-seq peaks. "length" - length of the peak. "value" - strength of the peak. "peak.log10pval", "peak.log10qval" - significance of the peak. "sequence" - sequence of the CTCF motif. "CTCF.log10pval", "CTCF.log10qval" - significance of the motif. "repeats" - whether the motif overlaps a repeat.