### 2022-05-25 EDA_SCREEN.Rmd
save to BED

---
title: "Analysis"
author: "Mikhail Dozmorov"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
```

# Libraries

```{r libraries}
library(tidyverse)
library(readxl)
library(writexl)
library(GenomicRanges)
library(rtracklayer)
```

# Settings

```{r settings}
# Project folder path
dir_data <- "/Users/mdozmorov/Documents/Work/GitHub/CTCF.dev"

# Input files
# Human CTCF-bound cCRE (hg38)
fileNameIn1 <- file.path(dir_data, "data/GRCh38-CTCF.bed")
# download.file("https://api.wenglab.org/screen_v13/fdownloads/cCREs/GRCh38-CTCF.bed", fileNameIn1)
# Mouse CTCF-bound cCRE (mm10)
fileNameIn2 <- file.path(dir_data, "data/mm10-CTCF.bed")
# download.file("https://api.wenglab.org/screen_v13/fdownloads/cCREs/mm10-CTCF.bed", fileNameIn2)
# Output files
fileNameOut1 <- file.path(dir_data, "results/hg38.SCREEN.GRCh38_CTCF.RData")
fileNameOut1.1 <- file.path(dir_data, "results/hg38.SCREEN.GRCh38_CTCF.bed")
fileNameOut2 <- file.path(dir_data, "results/mm10.SCREEN.mm10_CTCF.RData")
fileNameOut2.1 <- file.path(dir_data, "results/mm10.SCREEN.mm10_CTCF.bed")
```

PLS/pELS/dELS - Promoter-Like Signatures, proximal/distal Enhancer-Like Signatures

# Load hg38 data

```{r data}
# Read in data
mtx_human <- read_tsv(fileNameIn1, col_names = FALSE)
mtx_mouse <- read_tsv(fileNameIn2, col_names = FALSE)
```

## EDA

```{r}
table(mtx_human$X6) / length(mtx_human$X6)
table(mtx_mouse$X6) / length(mtx_mouse$X6)
```

## GRanges

```{r}
mtx_to_gr <- function(mtx = mtx_human, genome_id = "hg38") {
  gr <- GRanges(seqnames = mtx$X1, 
              IRanges(start = mtx$X2, 
                      end = mtx$X3))

  gr$ID1 <- mtx$X4
  gr$ID2 <- mtx$X5
  gr$Type <- mtx$X6

  # Get chromosome info and match it to the chromosome order in ctcfBED
  chrom_data <- GenomeInfoDb::getChromInfoFromUCSC(genome = genome_id)
  # Subset to common chromosomes
  common_chromosomes <- intersect(chrom_data$chrom, seqlevels(gr))
  chrom_data <- chrom_data[chrom_data$chrom %in% common_chromosomes, ]
  gr <- keepSeqlevels(gr, common_chromosomes, pruning.mode = "tidy")
  # Match order
  chrom_data <- chrom_data[match(seqlevels(gr), chrom_data$chrom), ]
  # Check if chromosome order is the same
  if (!all.equal(seqlevels(gr), chrom_data$chrom)) {
    print(paste("Chromosome order does not match for", genome_id, "genome."))
    break
  }
  # Assign seqinfo data
  seqlengths(gr) <- chrom_data$size
  isCircular(gr) <- chrom_data$circular
  genome(gr)     <- genome_id

  return(gr)
}

hg38.SCREEN.GRCh38_CTCF <- mtx_to_gr(mtx = mtx_human, genome_id = "hg38")
length(hg38.SCREEN.GRCh38_CTCF)
summary(width(hg38.SCREEN.GRCh38_CTCF))
mm10.SCREEN.mm10_CTCF <- mtx_to_gr(mtx = mtx_mouse, genome_id = "mm10")
length(mm10.SCREEN.mm10_CTCF)
summary(width(mm10.SCREEN.mm10_CTCF))

# Save as RData object. subdir is the character name of the CTCF GRanges variable
save(list = hg38.SCREEN.GRCh38_CTCF, file = fileNameOut1)
save(list = mm10.SCREEN.mm10_CTCF, file = fileNameOut2)
# load(file = paste0(subdir, ".RData"))
# Save BED files
export.bed(hg38.SCREEN.GRCh38_CTCF, fileNameOut1.1)
export.bed(mm10.SCREEN.mm10_CTCF, fileNameOut2.1)
```