### 2022-06-03 BED_to_BEDPE.Rmd
Add BED_to_BEDPE and FIMO_EDA

---
title: "Analysis"
author: "Mikhail Dozmorov"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
```

# Libraries

```{r libraries}
library(tidyverse)
library(readxl)
library(writexl)
library(cowplot)
library(stringr)
library("ggsci")
library(scales)
# scales::show_col(pal_lancet("lanonc")(8))
mycols = pal_lancet("lanonc")(8)
# Color palette for the heatmap, https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/colorPaletteCheatsheet.pdf
col3 <- colorRampPalette(c('blue', 'white', 'red'))(20)
# col3 <- colorRampPalette(c('blue', 'gray', 'yellow'))(20)
# col3 <- colorRampPalette(c('green', 'black', 'red'))(20)
# col3 <- colorRamps::green2red(n = 20)

```

# Settings

```{r settings}
# Project folder path
dir_data <- "/Users/mdozmorov/Documents/Data/GoogleDrive/Avocado_preciseTAD/Maggie/GM12878/PTBR_Peakachu_outputs/"

# Input files
fileNameIn1 <- file.path(dir_data, "")
# Output files
fileNameOut1 <- file.path(dir_data, "")
```

# Convert BED to BEDPE

```{r data}
# Read in data
files <- list.files(dir_data, pattern = "*.bed$")
for (fil in files) {
  print(fil)
  # BED data
  mtx <- read_tsv(file.path(dir_data, fil), col_names = FALSE)
  # Convert to pairs
  bedpe <- rbind()
  for (i in 1:(nrow(mtx) - 1)) {
    bedpe <-rbind(bedpe, c(mtx[i, 1:3], mtx[i + 1, 1:3]))
  }
  # BEDPE file name
  newfile <- file.path(dir_data, paste0(fil, "pe"))
  write.table(bedpe, newfile, quote = FALSE, sep = "\t", col.names = FALSE, row.names = FALSE)
  # write_tsv(as.data.frame(bedpe), newfile, col_names = FALSE)
}
```