### 2022-06-03 FIMO_EDA.Rmd

---
title: "Analysis"
author: "Mikhail Dozmorov"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
```

# Libraries

```{r libraries}
library(tidyverse)
library(readxl)
library(writexl)
library(cowplot)
library(stringr)
library("ggsci")
library(scales)
# scales::show_col(pal_lancet("lanonc")(8))
mycols = pal_lancet("lanonc")(8)
# Color palette for the heatmap, https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/colorPaletteCheatsheet.pdf
col3 <- colorRampPalette(c('blue', 'white', 'red'))(20)
# col3 <- colorRampPalette(c('blue', 'gray', 'yellow'))(20)
# col3 <- colorRampPalette(c('green', 'black', 'red'))(20)
# col3 <- colorRamps::green2red(n = 20)

```

# Settings

```{r settings}
dir_data <- "/Users/mdozmorov/Documents/Data/GoogleDrive/CTCF.dev/merlot"
# Motifs
motif_human  <- c("CIS_BP_2.00_Homo_sapiens", "CTCFBSDB_PWM", "HOCOMOCOv11_core_HUMAN_mono_meme_format", "JASPAR2022_CORE_vertebrates_non_redundant_v2", "jolma2013", "SwissRegulon_human_and_mouse")
motif_mouse  <- c("CIS_BP_2.00_Mus_musculus", "CTCFBSDB_PWM", "HOCOMOCOv11_core_MOUSE_mono_meme_format", "JASPAR2022_CORE_vertebrates_non_redundant_v2", "jolma2013", "SwissRegulon_human_and_mouse")
# Genomes
genome_human <- c("T2T", "hg38", "hg19")
genome_mouse <- c("mm10", "mm9")
```

```{r hg18, eval=FALSE}
# Human data hg18
pvals <- rbind()
pval_threshold <- 1e-6
for (gen in c("hg18", genome_human)) {
  for (mot in "JASPAR2022_MA0139.1") {
    print(paste(gen, mot, sep = "."))
    mtx <- read_tsv(fileNameIn, col_names = FALSE, col_types = c("ciicdciddc"))
    fileNameIn <- file.path(dir_data, paste(gen, mot, "bed", sep = "."))
    print(paste("Number of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold), digits = 5) ))
    print(paste("Proportion of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold) / nrow(mtx), digits = 5) ))
    pvals <- rbind(pvals, cbind(-log10(mtx$X8), rep(gen, nrow(mtx))))
    # pvals$pvals <- c(pvals$pvals, -log10(mtx$X8))
    # pvals$assembly <- c(pvals$assembly, rep(gen, nrow(mtx)))
  }
}
pvals <- as.data.frame(pvals)
pvals[, 1] <- as.numeric(pvals[, 1])
colnames(pvals) <- c("-log10(p-value)", "Assembly")

ggplot(pvals, aes(x = `-log10(p-value)`, color = Assembly)) +
  geom_density()
```

# Human data all

```{r hgall}
pvals <- rbind()
pval_threshold <- 1e-6
for (gen in "hg38") {
  for (mot in motif_human) {
    print(paste(gen, mot, sep = "."))
    fileNameIn <- file.path(dir_data, paste(gen, mot, "bed", sep = "."))
    mtx <- read_tsv(fileNameIn, col_names = FALSE, col_types = c("ciicdciddc"))
    print(paste("Number of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold), digits = 5) ))
    print(paste("Proportion of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold) / nrow(mtx), digits = 5) ))
    pvals <- rbind(pvals, cbind(-log10(mtx$X8), rep(mot, nrow(mtx))))
  }
}
pvals <- as.data.frame(pvals)
pvals[, 1] <- as.numeric(pvals[, 1])
colnames(pvals) <- c("-log10(p-value)", "MotifDB")

ggplot(pvals, aes(x = `-log10(p-value)`, color = MotifDB)) +
  geom_density() +
  geom_vline(xintercept = -log10(pval_threshold), linetype = "dashed", color = "red") +
  scale_color_manual(values = mycols[1:6]) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      # axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
      plot.title = element_text(hjust = 0.5)) +
  ylab("Density")

fileNameOut1 <- file.path(dir_data, "Figure_human_pvalues.svg")
ggsave(filename = fileNameOut1, width = 7, height = 3)
```

# Mouse data all

```{r mmall}
pvals <- rbind()
pval_threshold <- 1e-6
for (gen in "mm10") {
  for (mot in motif_mouse) {
    print(paste(gen, mot, sep = "."))
    fileNameIn <- file.path(dir_data, paste(gen, mot, "bed", sep = "."))
    mtx <- read_tsv(fileNameIn, col_names = FALSE, col_types = c("ciicdciddc"))
    print(paste("Number of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold), digits = 5) ))
    print(paste("Proportion of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold) / nrow(mtx), digits = 5) ))
    pvals <- rbind(pvals, cbind(-log10(mtx$X8), rep(mot, nrow(mtx))))
  }
}
pvals <- as.data.frame(pvals)
pvals[, 1] <- as.numeric(pvals[, 1])
colnames(pvals) <- c("-log10(p-value)", "MotifDB")

ggplot(pvals, aes(x = `-log10(p-value)`, color = MotifDB)) +
  geom_density() +
  geom_vline(xintercept = -log10(pval_threshold), linetype = "dashed", color = "red") +
  scale_color_manual(values = mycols[1:6]) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      # axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
      plot.title = element_text(hjust = 0.5)) +
  ylab("Density")

fileNameOut1 <- file.path(dir_data, "Figure_mouse_pvalues.svg")
ggsave(filename = fileNameOut1, width = 7, height = 3)
```