### 2022-06-03 FIMO_processing.Rmd
Savind extended summary and tweaks

---
title: "Analysis"
author: "Mikhail Dozmorov"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
```

# Libraries

```{r libraries}
library(tidyverse)
library(readxl)
library(writexl)
library(cowplot)
library(stringr)
library("ggsci")
library(scales)
# scales::show_col(pal_lancet("lanonc")(8))
mycols = pal_lancet("lanonc")(8)
library(GenomicRanges)
library(rtracklayer)
```

# Settings

```{r settings}
# Project folder path
dir_data <- "/Users/mdozmorov/Documents/Data/GoogleDrive/CTCF.dev/merlot"
# Motifs
motif_human  <- c("CIS_BP_2.00_Homo_sapiens", "CTCFBSDB_PWM", "HOCOMOCOv11_core_HUMAN_mono_meme_format", "JASPAR2022_CORE_vertebrates_non_redundant_v2", "jolma2013", "SwissRegulon_human_and_mouse")
motif_mouse  <- c("CIS_BP_2.00_Mus_musculus", "CTCFBSDB_PWM", "HOCOMOCOv11_core_MOUSE_mono_meme_format", "JASPAR2022_CORE_vertebrates_non_redundant_v2", "jolma2013", "SwissRegulon_human_and_mouse")
# Genomes
genome_human <- c("T2T", "hg38", "hg19")
genome_mouse <- c("mm10", "mm9")
# Chromosomes
chromosome_human <- c("chr1", "chr2", "chr3", "chr4", "chr5", "chr6", "chr7", "chr8", "chr9", "chr10", "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20", "chr21", "chr22", "chrX", "chrY")
chromosome_mouse <- c("chr1", "chr2", "chr3", "chr4", "chr5", "chr6", "chr7", "chr8", "chr9", "chr10", "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chrX", "chrY")

# Input files
fileNameIn1 <- file.path(dir_data, "")
```

# Process data

```{r data}
combine_fimo <- function(motif = "SwissRegulon_human_and_mouse", genome_id = "T2T", chromosomes = chromosome_human) {
  ctcfGR <- c()
  for(chrom in chromosomes) {
    # print(chrom)
    fileNameIn <- file.path(dir_data, paste(motif, genome_id, chrom, sep = "_"), "fimo.txt.gz")
    # Read "fimo.bed" created by "fimo.qsub"
    ctcfBED <- read_tsv(fileNameIn, col_types = c("ccddcdddc"))
    # Convert to GRanges object
    ctcf_chr_gr <- GRanges(seqnames = ctcfBED$`sequence name`, 
                      IRanges(start = ctcfBED$start, end = ctcfBED$stop),
                      strand = ctcfBED$strand)
    # Add metadata
    ctcf_chr_gr$name = ctcfBED$`#pattern name`
    ctcf_chr_gr$score = ctcfBED$score
    ctcf_chr_gr$pvalue = ctcfBED$`p-value`
    ctcf_chr_gr$qvalue = ctcfBED$`q-value`
    ctcf_chr_gr$sequence = ctcfBED$`matched sequence`
    ctcfGR <- c(ctcfGR, ctcf_chr_gr)
  }
  # Combine all chromosomes
  ctcfGR <- do.call(c, as(ctcfGR, "GRangesList"))
  # Sort
  ctcfGR <- ctcfGR %>% sort()

  # Add seqinfo
  # Parse out genome ID from the folder name, to get hg19, hg38, mm9, or mm10
  if (genome_id == "T2T") {
    # Seqinfor for T2T genome
    chrom_data <- GenomeInfoDb::getChromInfoFromNCBI(assembly = "GCA_009914755.4")
    chrom_data$AssignedMolecule <- as.character(paste0("chr", chrom_data$AssignedMolecule))
    chrom_data <- chrom_data[chrom_data$AssignedMolecule %in% seqlevels(ctcfGR), ]
    chrom_data <- chrom_data[match(seqlevels(ctcfGR), chrom_data$AssignedMolecule), ]
    # Check if chromosome order is the same
    if (!all.equal(seqlevels(ctcfGR), chrom_data$AssignedMolecule)) {
      print(paste("Chromosome order does not match for", genome_id, "genome."))
      break
    }
    # Assign seqinfo data
    seqlengths(ctcfGR) <- chrom_data$SequenceLength
    isCircular(ctcfGR) <- ifelse(is.na(chrom_data$circular), FALSE, TRUE)
    genome(ctcfGR)     <- "GCA_009914755.4"
  } else {
    # Get chromosome info and match it to the chromosome order in ctcfBED
    chrom_data <- GenomeInfoDb::getChromInfoFromUCSC(genome = genome_id)
    chrom_data <- chrom_data[chrom_data$chrom %in% seqlevels(ctcfGR), ]
    chrom_data <- chrom_data[match(seqlevels(ctcfGR), chrom_data$chrom), ]
    # Check if chromosome order is the same
    if (!all.equal(seqlevels(ctcfGR), chrom_data$chrom)) {
      print(paste("Chromosome order does not match for", genome_id, "genome."))
      break
    }
    # Assign seqinfo data
    seqlengths(ctcfGR) <- chrom_data$size
    isCircular(ctcfGR) <- chrom_data$circular
    genome(ctcfGR)     <- genome_id
  }

  ctcfGR_to_save <- paste(genome_id, motif, sep = ".")
  fileNameOut1 <- file.path(dir_data, paste0(ctcfGR_to_save, ".RData"))
  fileNameOut1.1 <- file.path(dir_data, paste0(ctcfGR_to_save, ".bed"))
  # Assign this object to the subfolder-specific variable name
  assign(ctcfGR_to_save, ctcfGR)
  # Save as RData object. ctcfGR_to_save is the character name of the CTCF GRanges variable
  save(list = ctcfGR_to_save, file = fileNameOut1)
  # load(file = fileNameOut1)
  # export.bed(eval(parse(text = ctcfGR_to_save)), fileNameOut1.1)
  ctcfDF_to_save <- as.data.frame(eval(parse(text = ctcfGR_to_save)))
  ctcfDF_to_save <- ctcfDF_to_save[, c("seqnames", "start", "end", "name", "score", "strand", "width", "pvalue", "qvalue", "sequence")]
  write_tsv(ctcfDF_to_save, fileNameOut1.1, col_names = FALSE)
}
```

# Human data hg18

```{r hg18}
for (gen in c("hg18", genome_human)) {
  for (mot in "JASPAR2022_MA0139.1") {
    print(paste(gen, mot, sep = "."))
    combine_fimo(motif = mot, genome_id = gen, chromosomes = chromosome_human)
  }
}
```

# Human data all

```{r hgall}
# motif_human  <- c("CTCFBSDB_PWM_corrected", "HOCOMOCOv11_core_HUMAN_mono_meme_format", "JASPAR2022_CORE_vertebrates_non-redundant_v2", "jolma2013_corrected", "SwissRegulon_human_and_mouse")
for (gen in genome_human) {
  for (mot in motif_human) {
    print(paste(gen, mot, sep = "."))
    combine_fimo(motif = mot, genome_id = gen, chromosomes = chromosome_human)
  }
}
```

# Mouse data all

```{r mmall}
for (gen in genome_mouse) {
  for (mot in motif_mouse) {
    print(paste(gen, mot, sep = "."))
    combine_fimo(motif = mot, genome_id = gen, chromosomes = chromosome_mouse)
  }
}
```

# Summary

```{r}
# for file in `find . -type f -name "*MA0139*"`; do
#   echo $file
#   newfile=`echo $file | sed 's/MA0139/JASPAR2022_MA0139/'`
#   echo $newfile
#   mv $file $newfile
# done
# 
# Output files
fileNameOut1 <- file.path(dir_data, "log_PWMs.csv")
if (file.exists(fileNameOut1)) { file.remove(fileNameOut1) }
# Defined from FIMO_EDA
pval_threshold <- 1e-6
files <- list.files(dir_data, pattern = "*.bed", full.names = TRUE)
for (fil in files) {
  print(basename(fil) %>% str_remove(., ".bed"))
  mtx <- read_tsv(fil, col_names = FALSE, col_types = c("ciicdciddc"))
  colnames(mtx) <- c("seqnames", "start", "end", "name", "score", "strand", "width", "pvalue", "qvalue", "sequence")
  gr <- makeGRangesFromDataFrame(mtx, keep.extra.columns = TRUE)
  # Summary
  summary_to_save <- data.frame(File = basename(fil) %>% str_remove(., ".bed"),
                                Assembly = strsplit(basename(fil), ".", fixed = TRUE)[[1]][1],
                                `All (p-value threshold 1e-4)` = length(gr),
                                `Reduced (p-value threshold 1e-4)` = length(reduce(gr)),
                                `All (p-value threshold 1e-6)` = length(gr[gr$pvalue < pval_threshold]),
                                `Reduced (p-value threshold 1e-6)` = length(reduce(gr[gr$pvalue < pval_threshold])) )
  write_csv(summary_to_save, fileNameOut1, append = TRUE)
  gc()
}
```