### 2022-06-04 Logos.Rmd
Logo cloud for CIS-BP human

---
title: "Analysis"
author: "Mikhail Dozmorov"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
```

# Libraries

```{r libraries}
library(tidyverse)
library(readxl)
library(writexl)
library(cowplot)
library(stringr)
library("ggsci")
library(scales)
# scales::show_col(pal_lancet("lanonc")(8))
mycols = pal_lancet("lanonc")(8)
library(motifStack)
library(universalmotif)
```

# Settings

```{r settings}
# Project folder path
dir_data <- "/Users/mdozmorov/Documents/Data/GoogleDrive/CTCF.dev/"
```

```{r}
# Object to collect motifs
motifs_all <- list()
```

# JASPAR 2022

```{r fig.width = 6, fig.height = 4}
# Input files
fileNameIn1 <- file.path(dir_data, "motif_databases/JASPAR2022_CORE_vertebrates_non-redundant_v2.meme")
# Output files
fileNameOut1 <- file.path(dir_data, "motif_databases/JASPAR2022_CORE_vertebrates_non-redundant_v2.svg")
# Read in data
motifs <- importMatrix(fileNameIn1)
# plot(motifs[[1]])
motifStack(motifs)
dev.off()
svg(fileNameOut1, width = 6, height = 4)
motifStack(motifs)
dev.off()
# Standardize names
names(motifs) <- paste("JASPAR2022", "CORE", "vertebrates", "CTCF", names(motifs) %>% str_remove(., ".CTCF"), sep = "_")
# Append to the whole list
motifs_all <- c(motifs_all, motifs)
```

# HOCOMOCO v11

```{r fig.width = 6, fig.height = 2}
# Input files
fileNameIn1 <- file.path(dir_data, "motif_databases/HOCOMOCOv11_core_HUMAN_mono_meme_format.meme")
# Output files
fileNameOut1 <- file.path(dir_data, "motif_databases/HOCOMOCOv11_core_HUMAN_mono_meme_format.svg")
# Read in data
motifs <- importMatrix(fileNameIn1)
plot(motifs[[1]])
dev.off()
svg(fileNameOut1, width = 6, height = 2)
plot(motifs[[1]])
dev.off()
# Standardize names
names(motifs) <- paste("HOCOMOCO_v11", "core", "HUMAN", "CTCF", names(motifs) %>% str_remove(., "CTCF_HUMAN."), sep = "_")
# Append to the whole list
motifs_all <- c(motifs_all, motifs)
```

```{r fig.width = 6, fig.height = 2}
# Input files
fileNameIn1 <- file.path(dir_data, "motif_databases/HOCOMOCOv11_core_MOUSE_mono_meme_format.meme")
# Output files
fileNameOut1 <- file.path(dir_data, "motif_databases/HOCOMOCOv11_core_MOUSE_mono_meme_format.svg")
# Read in data
motifs <- importMatrix(fileNameIn1)
plot(motifs[[1]])
dev.off()
svg(fileNameOut1, width = 6, height = 2)
plot(motifs[[1]])
dev.off()
# Standardize names
names(motifs) <- paste("HOCOMOCO_v11", "core", "MOUSE", "CTCF", names(motifs) %>% str_remove(., "CTCF_MOUSE."), sep = "_")
# Append to the whole list
motifs_all <- c(motifs_all, motifs)
```

# SwissRegulon

```{r fig.width = 6, fig.height = 2}
# Input files
fileNameIn1 <- file.path(dir_data, "motif_databases/SwissRegulon_human_and_mouse.meme")
# Output files
fileNameOut1 <- file.path(dir_data, "motif_databases/SwissRegulon_human_and_mouse.svg")
# Read in data
motifs <- importMatrix(fileNameIn1)
plot(motifs[[1]])
dev.off()
svg(fileNameOut1, width = 6, height = 2)
plot(motifs[[1]])
dev.off()
# Standardize names
names(motifs) <- paste("SwissRegulon", "HUMAN_MOUSE", "CTCF", names(motifs) %>% str_remove(., "CTCF.") %>% str_remove(., fixed(".")), sep = "_")
# Append to the whole list
motifs_all <- c(motifs_all, motifs)
```

# Jomla 2013

Needs to be reverse complement

```{r fig.width = 6, fig.height = 2}
# Input files
fileNameIn1 <- file.path(dir_data, "motif_databases/jolma2013.meme")
# Output files
fileNameOut1 <- file.path(dir_data, "motif_databases/jolma2013.svg")
# Reverse complement
fileNameOut2 <- file.path(dir_data, "motif_databases/jolma2013_RC.meme")
# Read in data
motifs <- importMatrix(fileNameIn1)
# Reverse complement
motifs_RC <- motif_rc(motifs[[1]])
plot(motifs_RC)
dev.off()
svg(fileNameOut1, width = 6, height = 2)
plot(motifs[[1]])
dev.off()
# Save Reverse complement
write_meme(motifs_RC, fileNameOut2)
# Standardize names
motifs_RC@name <- names(motifs) %>% str_remove(., fixed(".")) # paste("Jomla2013", "HUMAN_MOUSE", "HT_SELEX", "CTCF", names(motifs) %>% str_remove(., "CTCF_") %>% str_remove(., fixed(".")), sep = "_")
# Append to the whole list
motifs_all <- c(motifs_all, motifs_RC)
```

# Clustering

```{r}
## plot stacks with hierarchical tree
motifStack(motifs_all, layout="tree") 
# Output files
fileNameOut1 <- file.path(dir_data, "motif_databases/Figure_clustered_main_PWMs.svg")
dev.off()
svg(fileNameOut1, width = 8, height = 8)
motifStack(motifs_all, layout="tree") 
dev.off()
```

# CIS_BP Homo Sapiens

```{r}
# Modify input dir
dir_data1 <- file.path(dir_data, "Homo_sapiens_2022_05_20_3_16_pm/pwms_CTCF_motifs")
files <- list.files(dir_data1, pattern = "*.txt", full.names = TRUE)
# Output file
fileNameOut1 <- file.path(dir_data1, "CIS-BP_2.00_Homo_sapiens.meme")
fileNameOut2 <- file.path(dir_data1, "CIS-BP_2.00_Homo_sapiens.svg")
# Standard header
write_lines("MEME version 5.4.1 (Tue Mar 1 19:18:48 2022 -0800)", file = fileNameOut1)
write_lines("ALPHABET= ACGT", file = fileNameOut1, append = TRUE)
write_lines("strands: + -", file = fileNameOut1, append = TRUE)
write_lines("Background letter frequencies (from uniform background):", file = fileNameOut1, append = TRUE)
write_lines("A 0.25000 C 0.25000 G 0.25000 T 0.25000 ", file = fileNameOut1, append = TRUE)
# Store number of bases
nsites <- c()
for(i in 1:length(files)) {
  file=files[i]
  # Make motif name
  motif_name <- basename(file) %>% str_remove(., ".txt") %>% str_replace(., ".", "_")
  mtx <- read_tsv(file, col_types = c("ccccc"))
  # If non-empty, proceed
  if (nrow(mtx) > 0) {
    # Collect number of bases
    nsites <- c(nsites, nrow(mtx))
    # Motif name
    write_lines(paste0("MOTIF ", motif_name), file = fileNameOut1, append = TRUE)
    # Header line
    header_line <- paste0("letter-probability matrix: alength= 4 w= ", nrow(mtx), " nsites= 20 E= 0")
    write_lines(header_line, file = fileNameOut1, append = TRUE)
    # Make first numerical column empty
    mtx$Pos <- ""
    # Write actual data
    write_tsv(mtx, file = fileNameOut1, col_names = FALSE, append = TRUE)
    # Separate from the next motif
    write_lines("", file = fileNameOut1, append = TRUE)
  }
}
print("Number of motifs")
print(length(nsites))
print("Summary of motif widths")
summary(nsites)
# Test if can be imported
motifs_all <- importMatrix(fileNameOut1)
if (!file.exists(fileNameOut2)) {
  dev.off()
  svg(fileNameOut2, width = 8, height = 4)
  # motifStack(motifs_all, layout="tree") 
  ## use matalign to calculate the distances of motifs
  hc <- clusterMotifs(motifs_all)
  ## convert the hclust to phylog object
  library(ade4)
  phylog1 <- ade4::hclust2phylog(hc)
  ## reorder the motifs_all by the order of hclust
  leaves <- names(phylog1$leaves)
  motifs_all <- motifs_all[leaves]
  ## extract the motif signatures
  motifSig <- motifSignature(motifs_all, phylog1, cutoffPval=0.0001, min.freq=1)
  ## draw the motifs with a tag-cloud style.
  motifCloud(motifSig, scale=c(1, 0.55), rot.per = 0,
             layout="rectangles", 
             freq.col = "black")
  dev.off()
}
```

# CIS_BP Mus Musculus

```{r}
# Modify input dir
dir_data1 <- file.path(dir_data, "Mus_musculus_2022_05_20_4_01_pm/pwms_CTCF_motifs")
files <- list.files(dir_data1, pattern = "*.txt", full.names = TRUE)
# Output file
fileNameOut1 <- file.path(dir_data1, "CIS-BP_2.00_Mus_musculus.meme")
fileNameOut2 <- file.path(dir_data1, "CIS-BP_2.00_Mus_musculus.svg")
# Standard header
write_lines("MEME version 5.4.1 (Tue Mar 1 19:18:48 2022 -0800)", file = fileNameOut1)
write_lines("ALPHABET= ACGT", file = fileNameOut1, append = TRUE)
write_lines("strands: + -", file = fileNameOut1, append = TRUE)
write_lines("Background letter frequencies (from uniform background):", file = fileNameOut1, append = TRUE)
write_lines("A 0.25000 C 0.25000 G 0.25000 T 0.25000 ", file = fileNameOut1, append = TRUE)
# Store number of bases
nsites <- c()
for(i in 1:length(files)) {
  file=files[i]
  # Make motif name
  motif_name <- basename(file) %>% str_remove(., ".txt")
  mtx <- read_tsv(file, col_types = c("ccccc"))
  # If non-empty, proceed
  if (nrow(mtx) > 0) {
    # Collect number of bases
    nsites <- c(nsites, nrow(mtx))
    # Motif name
    write_lines(paste0("MOTIF ", motif_name), file = fileNameOut1, append = TRUE)
    # Header line
    header_line <- paste0("letter-probability matrix: alength= 4 w= ", nrow(mtx), " nsites= 20 E= 0")
    write_lines(header_line, file = fileNameOut1, append = TRUE)
    # Make first numerical column empty
    mtx$Pos <- ""
    # Write actual data
    write_tsv(mtx, file = fileNameOut1, col_names = FALSE, append = TRUE)
    # Separate from the next motif
    write_lines("", file = fileNameOut1, append = TRUE)
  }
}
print("Number of motifs")
print(length(nsites))
print("Summary of motif widths")
summary(nsites)
# Test if can be imported
motifs_all <- importMatrix(fileNameOut1)
if (!file.exists(fileNameOut2)) {
  dev.off()
  svg(fileNameOut2, width = 8, height = 2.5)
  motifStack(motifs_all, layout="tree") 
  dev.off()
}
```

# CTCFBSDB

EMBL_M1 needs to be reverse complement

```{r}
fileNameIn1 <- file.path(dir_data, "CTCFBSDB_PWM.meme")
# Reverse complement
fileNameOut1 <- file.path(dir_data, "CTCFBSDB_PWM_corrected.meme")
# Read in data
motifs <- importMatrix(fileNameIn1)
# Correct EMBL_M1
motifs[[1]] <- motif_rc(motifs[[1]])
# Save the corrected matrix
write_meme(motifs, fileNameOut1)
# Plot
motifStack(motifs, layout="tree")
# Output files
fileNameOut1 <- file.path(dir_data, "motif_databases/Figure_clustered_CTCFBSDB_PWMs.svg")
dev.off()
svg(fileNameOut1, width = 8, height = 8)
motifStack(motifs, layout="tree") 
dev.off()
```

# ENCODE

```{r eval=FALSE}
# Modify input dir
dir_data1 <- file.path(dir_data, "ENCODE_motifs_CTCF")
files <- list.files(dir_data1, pattern = "*.txt", full.names = TRUE)
# Output file
fileNameOut1 <- file.path(dir_data1, "motifs_CTCF.meme")
fileNameOut2 <- file.path(dir_data1, "motifs_CTCF.svg")
# Standard header
write_lines("MEME version 5.4.1 (Tue Mar 1 19:18:48 2022 -0800)", file = fileNameOut1)
write_lines("ALPHABET= ACGT", file = fileNameOut1, append = TRUE)
write_lines("strands: + -", file = fileNameOut1, append = TRUE)
write_lines("Background letter frequencies (from uniform background):", file = fileNameOut1, append = TRUE)
write_lines("A 0.25000 C 0.25000 G 0.25000 T 0.25000 ", file = fileNameOut1, append = TRUE)
# Store number of bases
nsites <- c()
for(i in 1:length(files)) {
  fileNameIn=files[i]
  mtx <- read.table(fileNameIn, sep = " ", fill = TRUE)
  # Make motif name
  motif_name <- strsplit(mtx[1, 1], "\t")[[1]]
  motif_id <- motif_name[1] %>% str_remove(">")
  motif_alt <- motif_name[length(motif_name)]
  # Drop header row
  mtx <- mtx[2:nrow(mtx), ]
  # If non-empty, proceed
  if (nrow(mtx) > 0) {
    # Collect number of bases
    nsites <- c(nsites, nrow(mtx))
    # Motif name
    write_lines(paste0("MOTIF ", paste(motif_id, motif_alt)), file = fileNameOut1, append = TRUE)
    # Header line
    header_line <- paste0("letter-probability matrix: alength= 4 w= ", nrow(mtx), " nsites= 20 E= 0")
    write_lines(header_line, file = fileNameOut1, append = TRUE)
    # Make first numerical column empty
    mtx[, 1] <- ""
    # Write actual data
    write_tsv(mtx, file = fileNameOut1, col_names = FALSE, append = TRUE)
    # Separate from the next motif
    write_lines("", file = fileNameOut1, append = TRUE)
  }
}
print("Number of motifs")
print(length(nsites))
print("Summary of motif widths")
summary(nsites)
# Test if can be imported
motifs_all <- importMatrix(fileNameOut1)
if (!file.exists(fileNameOut2)) {
  dev.off()
  svg(fileNameOut2, width = 10, height = 14)
  motifStack(motifs_all, layout="tree") 
  dev.off()
  }
```

```{r eval=FALSE}
# Modify input file
fileNameIn1 <- file.path(dir_data, "motifs_CTCF.txt")
# Output file
fileNameOut1 <- file.path(dir_data1, "motifs_CTCF.meme")
fileNameOut2 <- file.path(dir_data1, "motifs_CTCF.svg")
# Standard header
write_lines("MEME version 5.4.1 (Tue Mar 1 19:18:48 2022 -0800)", file = fileNameOut1)
write_lines("ALPHABET= ACGT", file = fileNameOut1, append = TRUE)
write_lines("strands: + -", file = fileNameOut1, append = TRUE)
write_lines("Background letter frequencies (from uniform background):", file = fileNameOut1, append = TRUE)
write_lines("A 0.25000 C 0.25000 G 0.25000 T 0.25000 ", file = fileNameOut1, append = TRUE)
# Store number of bases
nsites <- c()
for(i in 1:length(files)) {
  file=files[i]
  mtx <- read.table(fileNameIn1, sep = " ", fill = TRUE)
  # Make motif name
  motif_name <- strsplit(mtx[1, 1], "\t")[[1]]
  motif_id <- motif_name[1] %>% str_remove(">")
  motif_alt <- motif_name[length(motif_name)]
  # If non-empty, proceed
  if (nrow(mtx) > 0) {
    # Collect number of bases
    nsites <- c(nsites, nrow(mtx))
    # Motif name
    write_lines(paste0("MOTIF ", motif_name), file = fileNameOut1, append = TRUE)
    # Header line
    header_line <- paste0("letter-probability matrix: alength= 4 w= ", nrow(mtx), " nsites= 20 E= 0")
    write_lines(header_line, file = fileNameOut1, append = TRUE)
    # Make first numerical column empty
    mtx$Pos <- ""
    # Write actual data
    write_tsv(mtx, file = fileNameOut1, col_names = FALSE, append = TRUE)
    # Separate from the next motif
    write_lines("", file = fileNameOut1, append = TRUE)
  }
}
print("Number of motifs")
print(length(nsites))
print("Summary of motif widths")
summary(nsites)
# Test if can be imported
motifs_all <- importMatrix(fileNameOut1)
if (!file.exists(fileNameOut2)) {
  dev.off()
  svg(fileNameOut2, width = 8, height = 2.5)
  motifStack(motifs_all, layout="tree") 
  dev.off()
}
```