### 2022-06-06 EDA_liftOver_mm.Rmd
mm liftOver

---
title: "Analysis"
author: "Mikhail Dozmorov"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
```

# Libraries

```{r libraries}
library(tidyverse)
library(readxl)
library(writexl)
library(cowplot)
library(stringr)
library("ggsci")
library(scales)
# scales::show_col(pal_lancet("lanonc")(8))
mycols = pal_lancet("lanonc")(8)
library(pheatmap)
# Color palette for the heatmap, https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/colorPaletteCheatsheet.pdf
col3 <- colorRampPalette(c('blue', 'white', 'red'))(20)
# col3 <- colorRampPalette(c('blue', 'gray', 'yellow'))(20)
# col3 <- colorRampPalette(c('green', 'black', 'red'))(20)
# col3 <- colorRamps::green2red(n = 20)
library(rtracklayer)
```

# Settings

```{r settings}
# Project folder path
dir_data <- "/Users/mdozmorov/Documents/Data/GoogleDrive/CTCF.dev/CTCF_liftover/"

# Input files
fileNameIn1 <- file.path(dir_data, "sample_annotation_mm.csv")
# Output files
fileNameOut1 <- file.path(dir_data, "Figure_liftOverJaccard_mm.svg")
```

# Load data

```{r data}
sample_annotation <- read_csv(fileNameIn1)
# Read in data
files <- sample_annotation$File
genome_assembly <- sample_annotation$Assembly
# Object to store GRanges 
bed_list_all <- list() # All chromosomes
bed_list_standard <- list() # Standard chromosomes
# Additional columns
# extraCols_names <- c("character", "character", "character")
# names(extraCols_names) <- c("motif", "pval", "qval")
# Process each file
for(i in 1:length(files)) {
  print(basename(files[i]))
  # Read data
  bed_file <- import(file.path(dir_data, "data", files[i]), format = "bed") # , genome = genome_assembly[i], colnames = c("chrom", "start", "end", "name", "score", "strand", "motif", "pval", "qval"), extraCols = extraCols_names
  # Save all data
  bed_list_all <- c(bed_list_all, list(bed_file))
  # Keep standard chromosomes
  bed_file_standard <- keepStandardChromosomes(bed_file, pruning.mode="coarse")
  # Save filtered data
  bed_list_standard <- c(bed_list_standard, list(bed_file_standard))
}
# Add names
names(bed_list_all) <- sample_annotation$Name
names(bed_list_standard) <- sample_annotation$Name
```

## Jaccard / GenometriCorr

```{r jaccard, warning=FALSE, fig.width=6.5, fig.height=6}
save_pheatmap_svg <- function(x, filename, width=4.5, height=3, units = "in", res = 300) {
  stopifnot(!missing(x))
  stopifnot(!missing(filename))
  svg(filename, width=width, height=height)
  grid::grid.newpage()
  grid::grid.draw(x$gtable)
  dev.off()
}
# Jaccard calculations
jaccard <- function(gr_a, gr_b) {
  intersects <- GenomicRanges::intersect(gr_a, gr_b, ignore.strand = TRUE)
  intersection <- sum(width(intersects))
  union <- sum(width(GenomicRanges::union(gr_a, gr_b, ignore.strand = TRUE)))
  DataFrame(intersection, union, 
            jaccard = intersection/union,
             n_intersections = length(intersects))
}
# GenometriCorr of smaller query vs. larger reference
# Returns either Jaccard or ECDF area correlation
GCorr <- function(gr1, gr2, return_value = "relative.distances.ecdf.area.correlation") {
  # Select smaller query and larger reference
  # GenometriCorr itselv
  gr1_vs_gr2 <- GenometriCorrelation(query, reference, awhole.only = TRUE, showProgressBar = FALSE, permut.number = 0)
  # What to return
  if (return_value == "relative.distances.ecdf.area.correlation") {
    return(gr1_vs_gr2$awhole$relative.distances.ecdf.area.correlation)
  }
  if (return_value == "jaccard.measure") {
    return(gr1_vs_gr2$awhole$jaccard.measure)
  }
}

# Correlation matrix, empty
mtx_to_plot <- matrix(data = 0, nrow = length(bed_list_standard), ncol = length(bed_list_standard))
# Fill it in
for (i in 1:length(bed_list_standard)) {
  for (j in 1:length(bed_list_standard)) {
    # If diagonal, set to zero
    if (i == j) mtx_to_plot[i, j] <- 1
    # Process only one half, the other is symmetric
    if (i > j) {
      gr1 <- bed_list_standard[[i]]
      gr2 <- bed_list_standard[[j]]
      # Jaccard
      query <- gr1
      reference <- gr2
      mtx_to_plot[i, j] <- mtx_to_plot[j, i] <- jaccard(query, reference)[["jaccard"]]
      # GenometriCorr
      # if (length(gr1) >= length(gr2)) {
      #   reference <- gr1
      #   query <- gr2
      # } else {
      #   reference <- gr2
      #   query <- gr1
      # }
      # mtx_to_plot[i, j] <- mtx_to_plot[j, i] <- GCorr(query, reference)
    }
  }
}

# Trim row/colnames
rownames(mtx_to_plot) <- colnames(mtx_to_plot) <- str_trunc(names(bed_list_standard), width = 25) 
```

# Pheatmap

```{r}
# Adjust clustering method
# if(genome_assembly == "hg38") {
#   clust_method <- "euclidean"
# }
# if(genome_assembly == "mm10") {
#   clust_method <- "correlation"
# }
clust_method <- "euclidean"
# Save the plot
col3 <- colorRampPalette(c('white', 'white','white','white','white','white','white','orangered'))(20)
# all.equal(colnames(mtx_to_plot), sample_annotation$Name)
mydf <- data.frame(Assembly = sample_annotation$Assembly)
rownames(mydf) <- sample_annotation$Name
# png("man/figures/excluderanges_hg38_jaccard.png", width = 1000, height = 900, res = 200)
p <- pheatmap(data.matrix(mtx_to_plot), cluster_cols = T, cluster_rows = T, color = col3,
         clustering_method = "ward.D",# "ward.D", 
         clustering_distance_rows = clust_method, 
         clustering_distance_cols = clust_method, 
         annotation_row = mydf, annotation_colors = list(Assembly = c(mm9 = mycols[2], mm10 = mycols[4])),
         treeheight_row = 40,
         treeheight_col = 0, 
         display_numbers = TRUE)

# mtx_to_barplot <- data.frame(hg19 = mtx_to_plot["hg19", "hg19 (from hg38)"],
#                              hg38 = mtx_to_plot["hg38", "hg38 (from hg19)"],
#                              T2T = mtx_to_plot["T2T (from hg19)", "T2T (from hg38)"])
# mtx_to_barplot <- mtx_to_barplot %>% pivot_longer(., cols = everything(), names_to = "Assembly", values_to = "Jaccard")
# mtx_to_barplot$Assembly <- c("hg19 vs.\nhg19 from hg38", "hg38 vs.\nhg38 from hg19", "T2T from hg19 vs.\nT2T from hg38")
# mtx_to_barplot$Assembly <- factor(mtx_to_barplot$Assembly, levels = rev(mtx_to_barplot$Assembly))
# ggplot(mtx_to_barplot, aes(x = Assembly, y = Jaccard, fill = Assembly)) +
#   geom_bar(stat = "identity") +
#   coord_flip() +
#   theme_bw(base_size = 15) +
#   # get rid of the grid 
#   theme(panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(), 
#         text = element_text(size = 15),
#         legend.position = "none") + # c(0.8, 0.30)
#   scale_fill_manual(values = mycols[1:3]) + # change colors 
#   xlab("Assembly") +
#   ylab("Jaccard")

save_pheatmap_svg(p, filename = fileNameOut1, width = 7.5, height = 5.5, units = "in", res = 300)
# dev.off()
```